---
title: "Applied Exercise 11"
author: "Group 11"
date: "02/03/2021"
output:
  html_document:
    toc: no
    df_print: paged
  word_document:
    toc: no
  pdf_document:
    toc: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 3. Applied Exercise 
## Q11: We will now try to predict per capita crime rate in the Boston data set.

## a) Try out some of the regression methods explored in this chapter, such as best subset selection, the lasso, ridge regression, and PCR. Present and discuss results for the approaches that you consider.


### Loading Boston Data
```{r}
library(MASS)
data(Boston)
```

### Exploring the Boston Dataset

This data frame contains the following columns:

|       crim: per capita crime rate by town.

|       zn: proportion of residential land zoned for lots over 25,000 sq.ft.

|       indus: proportion of non-retail business acres per town.

|       chas: Charles River dummy variable (= 1 if tract bounds river; 0 otherwise).

|       nox: nitrogen oxides concentration (parts per 10 million).

|       rm: average number of rooms per dwelling.

|       age: proportion of owner-occupied units built prior to 1940.

|       dis: weighted mean of distances to five Boston employment centres.

|       rad: index of accessibility to radial highways.

|       tax: full-value property-tax rate per $10,000.

|       ptratio: pupil-teacher ratio by town.

|       black: 1000(Bk-0.63)^2 where Bk is the proportion of blacks by town.

|       lstat: lower status of the population (percent).

|       medv: median value of owner-occupied homes in $1000s.

```{r}
head(Boston)
dim(Boston)
summary(Boston)
sum(is.na(Boston))
cor(Boston,Boston$crim)
```

### Ridge Regression
```{r,warning=FALSE,message=FALSE}
library(glmnet)
x <- model.matrix(crim~., Boston)[,-1]
y <- Boston$crim
grid <- 10^seq(10, -2, length = 100)

#train/test splitting
set.seed(1)
train <- sample(1:nrow(x), nrow(x)/2)
test <- (-train)
y.test <- y[test]

#fitting a model on the training set
ridge.mod <- glmnet(x[train,], y[train], alpha=0, lambda =grid, thresh =1e-12)

#cross validation
set.seed(1)
cv.out <- cv.glmnet(x[train,], y[train], alpha=0)
plot(cv.out)
(bestlam_r <- cv.out$lambda.min)

#evaluating its MSE on the test set
ridge.pred <- predict(ridge.mod, s=bestlam_r, newx=x[test,])
mean((ridge.pred - y.test)^2)
```

We found that cross validation selected a lambda equal to 0.5919159 and MSE was 40.92116.

### Lasso
```{r}
#fitting a model on the training set
lasso.mod <- glmnet(x[train,], y[train], alpha = 1, lambda=grid)

#cross validation
set.seed(1)
cv.out <- cv.glmnet(x[train ,], y[train], alpha=1)
plot(cv.out)
(bestlam <- cv.out$lambda.min)

#evaluating its MSE on the test set
lasso.pred <- predict(lasso.mod, s=bestlam, newx = x[test,])
mean(( lasso.pred -y.test)^2)
```

These results indicated that cross validation selected a lambda equal to 0.05148183 and MSE was 40.99066.


### PCR
```{r,warning=FALSE,message=FALSE}
#fitting a model on the training set and computing cross validation errors
library(pls)
set.seed(1)
pcr.fit <- pcr(crim~., data=Boston, subset = train, scale=TRUE, validation='CV')
validationplot(pcr.fit, val.type="MSEP")

#evaluating its MSE on the test set
pcr.pred <- predict(pcr.fit, x[test ,], ncomp=13)
mean((pcr.pred - y.test)^2)
```
Consequently, the lowest cross-validation error occurs when M=13 component were used (i.e. no dimension reduction), and the test MSE was 41.54639.


## b) Propose a model (or set of models) that seem to perform well on this data set, and justify your answer. Make sure that you are evaluating model performance using validation set error, crossvalidation, or some other reasonable alternative, as opposed to using training error.

According to the values of MSE, it was suggested that Ridge Regression was the best model though all models had almost the same MSE.


## c) Does your chosen model involve all of the features in the data set? Why or why not?
```{r}
out <- glmnet(x,y,alpha=0)
predict(out,type="coefficients",s=bestlam_r)[1:14,]
```
Yes, it does as shown above because ridge regression includes all predictors in the final model.
