---
title: "Team Project 1 Group11"
author: "Taylor Ann Applewhite, Olena Shumeiko, Yuri Takanabe, Nic Yabar"
date: "02/03/2021"
output:
  html_document:
    df_print: paged
---

```{r}
rm(list = ls())
library(ISLR)
library(pls)
set.seed(2)
x = model.matrix(Salary~., Hitters)[,-1]
y = Hitters$Salary
y<-na.omit(y)
train = sample(1:nrow(x), nrow(x)/2)
test = (-train)
y.test=y[test]
```
This is the set up for the modeling. We want to get rid of the na values first, and then we will set up the testing and training data.

```{r}
pcr.fit = pcr(Salary ~., data = Hitters, scale=T, validation='CV')
summary(pcr.fit)
```
validation="CV" causes pcr() to compute the ten-fold cross-validation error
for each possible value of M, the number of principal components used. The
resulting fit can be examined using summary().
```{r}
validationplot(pcr.fit, val.type = "MSEP")
```

plot the cross-validation scores using the validationplot() function. Using val.type="MSEP" will cause the cross-validation MSE to be plotted.
We see that the smallest cross-validation error occurs when M = 18 components
are used. This is barely fewer than M = 19, which amounts to
simply performing least squares, because when all of the components are
used in PCR no dimension reduction occurs.
cross-validation error is roughly the same when only one
component is included in the model. This suggests that a model that uses
just a small number of components might suffice
Briefly,
we can think of this as the amount of information about the predictors or
the response that is captured using M principal components. For example,
setting M = 1 only captures 38.31% of all the variance, or information, in
the predictors. In contrast, using M = 6 increases the value to 88.63%. If
we were to use all M = p = 19 components, this would increase to 100%.
```{r}
set.seed(1)
pcr.fit=pcr(Salary~., data=Hitters ,subset =train ,scale =TRUE ,
            validation ="CV")
validationplot(pcr.fit, val.type = "MSEP")
```

Now we find that the lowest cross-validation error occurs when M = 5
component are used. We compute the test MSE as follows
```{r}
pcr.pred = predict(pcr.fit, x[test,],ncomp = 5)
mean((pcr.pred -y.test)^2)
```
This test set MSE is competitive with the results obtained using ridge regression
and the lasso. However, as a result of the way PCR is implemented,
the final model is more difficult to interpret because it does not perform
any kind of variable selection or even directly produce coefficient estimates
```{r}
pcr.fit=pcr(y~x,scale =TRUE ,ncomp =5)
summary(pcr.fit)
```
Finally, we fit PCR on the full data set, using M = 5, the number of
components identified by cross-validation

```{r}
set.seed(1)
pls.fit = plsr(Salary~., data=Hitters, subset=train, scale=T, validation='CV')
summary(pls.fit)
validationplot(pls.fit , val.type="MSEP")
```

The lowest cross-validation error occurs when only M = 1 partial least
squares directions are used. We now evaluate the corresponding test set
MSE.
```{r}
pls.pred = predict(pls.fit, x[test,], ncomp = 2)
mean((pls.pred-y.test)^2)
```
The test MSE is comparable to, but slightly less than, the test MSE
obtained using PCR.
```{r}
pls.fit=plsr(Salary~., data=Hitters ,scale=TRUE ,ncomp =2)
summary(pls.fit)
```
Finally, we perform PLS using the full data set, using M = 1, the number
of components identified by cross-validation.
Notice that the percentage of variance in Salary that the one-component
PLS fit explains, 43.05%, is almost as much as that explained using the
final five-component model PCR fit, 44.90 %. This is because PCR only
attempts to maximize the amount of variance explained in the predictors,
while PLS searches for directions that explain variance in both the predictors
and the response


```{r}
rm(list=ls())
library(caret)
library(ISLR)
data("Hitters")
set.seed(2)
sum(is.na(Hitters))
Hitters<- na.omit(Hitters)
```
We are removing any NA's to clean up our data

```{r}
dividedata <- createDataPartition(Hitters$Salary, p=0.8, list = F)
train <- Hitters[dividedata,]
test <- Hitters[-dividedata,]
```
Here we are setting up our training and testing data using the createDataPartition command in the Caret package

```{r}
pcr.fit <- train(Salary~., method="pcr", data=train, scale=T, trControl = trainControl('cv', number=10), tuneLength=19)

```
An additional argument is scale = TRUE for standardizing the variables to make them comparable.

Caret uses cross-validation to automatically identify the optimal number of principal components (ncomp) to be incorporated in the model.

Here, we’ll test 20 different values of the tuning parameter ncomp. This is specified using the option tuneLength. The optimal number of principal components is selected so that the cross-validation error (RMSE) is minimized.

```{r}
summary(pcr.fit$finalModel)
plot(pcr.fit)
pcr.fit$bestTune
```

Here we are looking for the number of components that reduces our RMSE and it is 5.  The bestTune tells us that 5 principal components is the optimal amount.

```{r}
newpred <- predict(pcr.fit, newdata = test)
data.frame(
  MSE = (caret::RMSE(newpred, test$Salary)^2),
  Rsquare = caret::R2(newpred, test$Salary)
)
```

The MSE is higher than before when using the lab model

We will now do the same process for the Partial Least Squares method
```{r}
set.seed(1)
pls.fit <- train(Salary~., data=train, method='pls', scale=T, trControl = trainControl('cv', number = 10), tuneLength = 19)
```

We use the same notation as the pcr.fit model with the only change being "method = 'pls'" now


```{r}
summary(pls.fit$finalModel)
plot(pls.fit)
pls.fit$bestTune
```

Here we can see that the best amount of components in this case to reduce the MSE is 10
This model explains 49% of the variance in salary using 10 components

```{r}
predictions <- predict(pls.fit, newdata = test)
data.frame(
  MSE = (caret::RMSE(predictions, test$Salary)^2),
  Rsquare = caret::R2(predictions, test$Salary)
)
```
The MSE in the PLS method is slightly lower than the PCR model using the Caret package.  
Notice that the percentage of variance in Salary that the ten-component
PLS fit explains, 48.98%, is more than what was explained using the
final five-component model PCR fit, 39.95 %. This is because PCR only
attempts to maximize the amount of variance explained in the predictors,
while PLS searches for directions that explain variance in both the predictors
and the response

